## Course Project: Personal Movement 
Loading and reprocessing the data
```{r, echo=TRUE, results="hide", cache=TRUE}
md <- read.csv("activity.csv", header=TRUE)
library(ggplot2)
library(lattice)
library(dplyr)
md$date <- as.Date(as.character(md$date), "%Y-%m-%d")
```

####What is mean total number of steps taken per day?
```{r Calculation, echo=TRUE, cache=TRUE }
##Calculation of total, mean, and median
total <- with(md, tapply(steps, date, na.rm=TRUE, FUN=sum))
mean <- with(md, tapply(steps, date, na.rm=TRUE, FUN=mean))
median <- with(md, tapply(steps, date, na.rm=TRUE, FUN=median))

##Making the histogram and showing the mean per day
png(file="figure0.png")
hist(total, col="green", main="Total Number of Steps Taken")
dev.off()
```


The mean total number of steps taken per day is as follows: 
```{r mean, echo=TRUE, cache=FALSE}
mean
```
####What is the average daily activity pattern?
```{r Average, echo=TRUE, cache=TRUE}
##Calculation of  average daily activity pattern
ave <- with(md, aggregate(steps, list(interval), na.rm=TRUE, FUN=mean))

##Plotting the average
png(file="figure1.png")
plot(ave, col="brown", type="l", ylab="Average Steps", xlab="5 Minute Interval")
dev.off()
```


The peak or maximum of the average number of steps taken is on the interval between 800 and 1000. This is explained by the time of the interval which is morning where people are highly locomotive and active since they incur a large number of steps taken in various activities such as going to work or school, attending to daily errands, and/or having a morning exercise at the gym or parks.




####Imputing missing values
```{r total number of missing values, echo=TRUE, result="hide", cache=TRUE}
##Imputing Missing Values

	##Total Number of Rows with Missing Values
	missing <- sum(is.na(md))
```
The total number of rows with missing values is `r missing`.
```{r NA replacement, echo=TRUE, results="hide", cache=TRUE}
##Create New Data Set
	md2 <- md

##Fill in of the missing values with their mean
  library(zoo)
  md2[] <- lapply(md2, na.aggregate) 
	
```

```{r new results, echo=TRUE, cache=TRUE}
##Calculation of total, mean, and median
total2 <- with(md2, tapply(steps, date, FUN=sum))
mean2 <- with(md2, tapply(steps, date, FUN=mean))
median2 <- with(md2, tapply(steps, date, FUN=median))

##Plotting the total number of steps taken
png(file="figure2.png")
hist(total2, col="green", main="Total Number of Steps Taken")
dev.off()
```

The insertion of values in place of NA values has increased the total daily number of steps taken. The difference can be seen from the left side and the middle of the new histogram as the frequency has increased from 25 to 35 while the right side has little increase.  




####Are there differences in activity patterns between weekdays and weekends?
```{r differences, echo=TRUE, cache=TRUE}
##Create a new column for weekdays and weekends
	md2$Day <- weekdays(md2$date)

##Identification for each observation
	md2$Day[which(md2$Day == "Monday")] <- "Weekdays"
	md2$Day[which(md2$Day == "Tuesday")] <- "Weekdays"
	md2$Day[which(md2$Day == "Wednesday")] <- "Weekdays"
	md2$Day[which(md2$Day == "Thursday")] <- "Weekdays"
	md2$Day[which(md2$Day == "Friday")] <- "Weekdays"
	md2$Day[which(md2$Day == "Saturday")] <- "Weekends"
	md2$Day[which(md2$Day == "Sunday")] <- "Weekends"

##Subsetting the weekends and weekdays
	Day <- subset(md2, Day == "Weekdays")
	End <- subset(md2, Day == "Weekends")

##Calculating the mean of weekends and weekdays
	ave1 <- with(Day, aggregate(steps, list(interval), na.rm=TRUE, FUN=mean))
	ave2 <- with(End, aggregate(steps, list(interval), na.rm=TRUE, FUN=mean))

##Plotting the mean of weekends and weekdays
	png(file="figure3.png")
	par(mfrow=c(2,1), mar=c(4,4,2,1))
	plot(ave1, col="brown", type="l", ylab="Average Steps", xlab="5 Minute Interval")
	title(main="Weekdays")
	plot(ave2, col="brown", type="l", ylab="Average Steps", xlab="5 Minute Interval")
	title(main="Weekends")
	dev.off()
```


There is a relative difference between weekdays and weekends. People tend to walk or make steps during the morning (evidenced by the intervals between 800 and 1000 ). Also, the highest average steps taken in an interval during weekdays is 150 and 100 for the weekends. This is due to the fact that people tend to walk during the peak hours or rush hours such as going to work or school and/or having a morning exercise at the park or gym. 